---
title: "Titanic Survival"
author: "Foram Patel"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This analysis attempts to predict the probability for survival of the Titanic passengers. In order to do this, we will use the different features available about the passengers, and based on the relationship between the attributes and the survivors we try to formulate the chances of a passenger surviving.

In this analysis we are determining the relationship between the features and a passenger’s chance of survival, by plotting graphs and running some data manipulation operations on the dataset and finally we define a function which predicts whether a passenger will survive or not.

##Exploratory Data Analysis and Logic

```{r Titanic}
library(dplyr)
filePathTrain <- "C:/Users/patel/Downloads/DA/trainTitanic.csv"
dfTrain <- as.data.frame(read.csv(filePathTrain, header=TRUE))
head(dfTrain)
```
```{r Sex}
##Survival on basis of sex
table(dfTrain$Sex)
s1 <- filter(dfTrain, Survived==1)
table(s1$Sex)
library(ggplot2)
ggplot(data=dfTrain, aes(Survived, fill=Sex)) +
geom_bar()
```
```{r Age}
##Survival on basis of Age
summarise(dfTrain, meanage = mean(Age, na.rm=TRUE))
agem30 <- filter(dfTrain, Age >= 30, Survived==1)
print(paste("Age more than 30 :", nrow(agem30)))
agel30 <- filter(dfTrain, Age < 30, Survived==1)
print(paste("Age less than 30 :", nrow(agel30)))
agem15 <- filter(dfTrain, Age < 30 & Age >= 15 , Survived==1)
print(paste("Age more than 15 less than 30 :", nrow(agem15)))
agel15 <- filter(dfTrain, Age < 15, Survived==1)
print(paste("Age less than 15 :", nrow(agel15)))
agem55 <- filter(dfTrain, Age <= 55 & Age >=30 , Survived==1)
print(paste("Age more than 30 less than 55 :", nrow(agem55)))
agem55 <- filter(dfTrain, Age > 55, Survived==1)
print(paste("Age more than 55 :", nrow(agem55)))
agel10 <- filter(dfTrain, Age < 10, Survived==1)
print(paste("Age less than 10 :", nrow(agel10)))
agel5 <- filter(dfTrain, Age < 5, Survived==1)
print(paste("Age less than 5 :", nrow(agel5)))
agel55 <- filter(dfTrain, Age >= 30 & Age < 55, Survived==1, Sex=="male")
print(paste("Age less than 55 and more than 30 :", nrow(agel55)))

ggplot(data=dfTrain, aes(x=factor(Age), color=Survived)) +
  geom_bar(stat="count", width=0.7, fill="white")

p<-ggplot(dfTrain, aes(x=Age, fill=Survived, color=Survived)) +
  geom_histogram(position="identity", alpha=0.5)
p

```


```{r q2}
  #total people died
  
  #females not survived
  notsurvivingFemales <- filter(dfTrain, Survived == 0, Sex == "female")
  notsurvivingFemales
  
  #male not survived
  notsurvivingMales <- filter(dfTrain, Survived == 0, Sex == "male")
  notsurvivingMales
  
  
  
  #total survivals
  diedfemales<-nrow(notsurvivingFemales)
 
  print(paste("Total males died",diedfemales))
  
  diedmales<-nrow(notsurvivingMales)
  
  print(paste("Total females died",diedmales))
 
  print(paste("Total people died",diedmales+diedfemales))
  
  ```
  
```{r q3}
  
  #Died pclass
  class<-filter(dfTrain,Survived==0)
  y<-select(class, PassengerId,Pclass)
  
  table(y$Pclass)##most occuring 
  print(paste("People died from 3 class the most "))
  print(paste("and the no of people died in class 3 :327"))

```
5.
```{r q4}
  ##ploting a bar graph
  ##survived vs not survived
  
    
  ggplot(data = dfTrain,aes(x=Survived,fill=Sex)) + geom_bar()
  
```
6.
```{r q5}
   
    #classswiese survivals    

    ggplot(data = dfTrain,aes(x=Pclass,fill=Survived==1)) + geom_bar()
    #True = Survived
    #false = Not survived
 
p<-ggplot(dfTrain, aes(x=Age, fill=Survived, color=Survived)) +
  geom_histogram(position="identity", alpha=0.5)
p

```
7.
```{r q6}

  #histogram Age vs Survuived
  ggplot(dfTrain, aes(x=Age, fill=Survived==1)) +
  geom_histogram(binwidth = 3)
 
```
8.
```{r q7}
  #age and survived people

  age1 <- filter(dfTrain, Survived == 1)


  
  #selecting form age
  age2<-select(age1,Age,Survived)
  age3<-na.omit(age2)
   
  #finding avg age
  avgAge<-mean(age3$Age)
  print(paste("Avg age of the people who have survived",round(avgAge)))
  
  
  
  
  
 
  
```
9.
```{r q8}
  
  #fare vs survived
  ggplot(dfTrain, aes(x=Fare, fill=Survived==1)) +
  geom_histogram()


```
10.
```{r q9}
  #SibSp and Survived

  ggplot(dfTrain,aes(x=SibSp,fill = Survived == 1)) + geom_bar()
  dfTrain
  

```

12.

```{r q11}
  #Parch and Survived

  ggplot(dfTrain,aes(x=Parch,fill = Survived == 1)) + geom_bar()
  dfTrain


```
13.
```{r q12}
  
  sib_par<-filter(dfTrain)
  table(sib_par$SibSp)
 
  sib_par<-filter(dfTrain,Survived == 1)
  table(sib_par$SibSp)
  
  sib_par<-filter(dfTrain,Survived == 1)
  table(sib_par$Parch)
```  
```{r q13}
###Survival on the basis of Siblings and Parents onboard
summarise(dfTrain, meanFare = mean(Fare, na.rm=TRUE))
s1 <- filter(dfTrain, Survived==1)
sort(table(s1$Fare), TRUE)


s1 <- filter(dfTrain, Survived==1, SibSp == 0, Parch == 0)
sort(table(s1$Pclass), TRUE)
nrow(dfTrain$Survived)

s1 <- filter(dfTrain, Survived==1, SibSp == 0, Parch > 0)
sort(table(s1$Pclass), TRUE)

s1 <- filter(dfTrain, Survived==1, SibSp > 0, Parch > 0)
sort(table(s1$Pclass), TRUE)

s1 <- filter(dfTrain, Survived==1, SibSp > 0, Parch == 0)
sort(table(s1$Pclass), TRUE)
```

14.
```{r q14}
 #maleFemaleAgePclass

#FemaleAgeFclass
femaleAge0to15FclassTot <- filter(dfTrain, Sex=="female", Age <= 15, Pclass==1)
femaleAge0to15FclassSurvived <- filter(dfTrain,  Sex=="female", Age <= 15, Survived==1, Pclass==1)

print(paste("Percentage of female Survivors in Age range 0 to 15 & First class: ", (nrow(femaleAge0to15FclassSurvived)/nrow(femaleAge0to15FclassTot))*100))




femaleAge15to30FclassTot <- filter(dfTrain, Sex=="female", Age > 15 & Age <= 30, Pclass==1)
femaleAge15to30FclassSurvived <- filter(dfTrain,  Sex=="female", Age > 15 & Age <= 30, Survived==1, Pclass==1)

print(paste("Percentage of female Survivors in Age range 15 to 30: ", (nrow(femaleAge15to30FclassSurvived)/nrow(femaleAge15to30FclassTot))*100))




femaleAge30to45FclassTot <- filter(dfTrain, Sex=="female", Age > 30 & Age <= 45, Pclass==1)
femaleAge30to45FclassSurvived <- filter(dfTrain,  Sex=="female", Age > 30 & Age <= 45, Survived==1, Pclass==1)

print(paste("Percentage of female Survivors in Age range 30 to 45: ", (nrow(femaleAge30to45FclassSurvived)/nrow(femaleAge30to45FclassTot))*100))



femaleAge45to60FclassTot <- filter(dfTrain, Sex=="female", Age > 45 & Age <= 60, Pclass==1)
femaleAge45to60FclassSurvived <- filter(dfTrain,  Sex=="female", Age > 45 & Age <= 60, Survived==1, Pclass==1)

print(paste("Percentage of female Survivors in Age range 45 to 60: ", (nrow(femaleAge45to60FclassSurvived)/nrow(femaleAge45to60FclassTot))*100))




femaleAgeGT60FclassTot <- filter(dfTrain, Sex=="female", Age >= 60, Pclass==1)
femaleAgeGT60FclassSurvived <- filter(dfTrain,  Sex=="female", Age >= 60, Survived==1, Pclass==1)

print(paste("Percentage of female Survivors in Age range GT 60: ", (nrow(femaleAgeGT60FclassSurvived)/nrow(femaleAgeGT60FclassTot))*100))



#FemaleAgeSclass

femaleAge0to15SclassTot <- filter(dfTrain, Sex=="female", Age <= 15, Pclass==2)
femaleAge0to15SclassSurvived <- filter(dfTrain,  Sex=="female", Age <= 15, Survived==1, Pclass==2)

print(paste("Percentage of female Survivors in Age range 0 to 15 & Second class: ", (nrow(femaleAge0to15SclassTot)/nrow(femaleAge0to15SclassTot))*100))




femaleAge15to30SclassTot <- filter(dfTrain, Sex=="female", Age > 15 & Age <= 30, Pclass==2)
femaleAge15to30SclassSurvived <- filter(dfTrain,  Sex=="female", Age > 15 & Age <= 30, Survived==1, Pclass==2)

print(paste("Percentage of female Survivors in Age range 15 to 30 & second class: ", (nrow(femaleAge15to30SclassSurvived)/nrow(femaleAge15to30SclassTot))*100))




femaleAge30to45SclassTot <- filter(dfTrain, Sex=="female", Age > 30 & Age <= 45, Pclass==2)
femaleAge30to45SclassSurvived <- filter(dfTrain,  Sex=="female", Age > 30 & Age <= 45, Survived==1, Pclass==2)

print(paste("Percentage of female Survivors in Age range 30 to 45 & Second class: ", (nrow(femaleAge30to45SclassSurvived)/nrow(femaleAge30to45SclassTot))*100))



femaleAge45to60SclassTot <- filter(dfTrain, Sex=="female", Age > 45 & Age <= 60, Pclass==3)
femaleAge45to60SclassSurvived <- filter(dfTrain,  Sex=="female", Age > 45 & Age <= 60, Survived==1, Pclass==3)

print(paste("Percentage of female Survivors in Age range 45 to 60 & Second Class: ", (nrow(femaleAge45to60SclassSurvived)/nrow(femaleAge45to60SclassTot))*100))




femaleAgeGT60SclassTot <- filter(dfTrain, Sex=="female", Age >= 60, Pclass==2)
femaleAgeGT60SclassSurvived <- filter(dfTrain,  Sex=="female", Age >= 60, Survived==1, Pclass==2)

print(paste("Percentage of female Survivors in Age range GT 60 & Second class: ", (nrow(femaleAgeGT60SclassSurvived)/nrow(femaleAgeGT60SclassTot))*100))

#FemaleAgeTclass

femaleAge0to15TclassTot <- filter(dfTrain, Sex=="female", Age <= 15, Pclass==3)
femaleAge0to15TclassSurvived <- filter(dfTrain,  Sex=="female", Age <= 15, Survived==1, Pclass==3)

print(paste("Percentage of female Survivors in Age range 0 to 15 & Third class: ", (nrow(femaleAge0to15TclassSurvived)/nrow(femaleAge0to15TclassTot))*100))




femaleAge15to30TclassTot <- filter(dfTrain, Sex=="female", Age > 15 & Age <= 30, Pclass==3)
femaleAge15to30TclassSurvived <- filter(dfTrain,  Sex=="female", Age > 15 & Age <= 30, Survived==1, Pclass==3)

print(paste("Percentage of female Survivors in Age range 15 to 30 in Third class: ", (nrow(femaleAge15to30TclassSurvived)/nrow(femaleAge15to30TclassTot))*100))




femaleAge30to45TclassTot <- filter(dfTrain, Sex=="female", Age > 30 & Age <= 45, Pclass==3)
femaleAge30to45TclassSurvived <- filter(dfTrain,  Sex=="female", Age > 30 & Age <= 45, Survived==1, Pclass==3)

print(paste("Percentage of female Survivors in Age range 30 to 45 in Third class: ", (nrow(femaleAge30to45TclassSurvived)/nrow(femaleAge30to45TclassTot))*100))



femaleAge45to60TclassTot <- filter(dfTrain, Sex=="female", Age > 45 & Age <= 60, Pclass==3)
femaleAge45to60TclassSurvived <- filter(dfTrain,  Sex=="female", Age > 45 & Age <= 60, Survived==1, Pclass==3)

print(paste("Percentage of female Survivors in Age range 45 to 60 in Third class: ", (nrow(femaleAge45to60TclassSurvived)/nrow(femaleAge45to60TclassTot))*100))




femaleAgeGT60TclassTot <- filter(dfTrain, Sex=="female", Age >= 60, Pclass==3)
femaleAgeGT60TclassSurvived <- filter(dfTrain,  Sex=="female", Age >= 60, Survived==1, Pclass==3)

femaleAgeGT60TclassSurvived
femaleAgeGT60TclassTot

print(paste("Percentage of female Survivors in Age range GT 60 in third class: ", (nrow(femaleAgeGT60TclassSurvived)/nrow(femaleAgeGT60TclassTot))*100))


#male

#maleAgeFclass
maleAge0to15FclassTot <- filter(dfTrain, Sex=="male", Age <= 15, Pclass==1)
maleAge0to15FclassSurvived <- filter(dfTrain,  Sex=="male", Age <= 15, Survived==1, Pclass==1)

print(paste("Percentage of male Survivors in Age range 0 to 15 & First class: ", (nrow(maleAge0to15FclassSurvived)/nrow(maleAge0to15FclassTot))*100))




maleAge15to30FclassTot <- filter(dfTrain, Sex=="male", Age > 15 & Age <= 30, Pclass==1)
maleAge15to30FclassSurvived <- filter(dfTrain,  Sex=="male", Age > 15 & Age <= 30, Survived==1, Pclass==1)

print(paste("Percentage of male Survivors in Age range 15 to 30: ", (nrow(maleAge15to30FclassSurvived)/nrow(maleAge15to30FclassTot))*100))




maleAge30to45FclassTot <- filter(dfTrain, Sex=="male", Age > 30 & Age <= 45, Pclass==1)
maleAge30to45FclassSurvived <- filter(dfTrain,  Sex=="male", Age > 30 & Age <= 45, Survived==1, Pclass==1)

print(paste("Percentage of male Survivors in Age range 30 to 45: ", (nrow(maleAge30to45FclassSurvived)/nrow(maleAge30to45FclassTot))*100))



maleAge45to60FclassTot <- filter(dfTrain, Sex=="female", Age > 45 & Age <= 60, Pclass==1)
maleAge45to60FclassSurvived <- filter(dfTrain,  Sex=="female", Age > 45 & Age <= 60, Survived==1, Pclass==1)

print(paste("Percentage of male Survivors in Age range 45 to 60: ", (nrow(maleAge45to60FclassSurvived)/nrow(maleAge45to60FclassTot))*100))




maleAgeGT60FclassTot <- filter(dfTrain, Sex=="male", Age >= 60, Pclass==1)
maleAgeGT60FclassSurvived <- filter(dfTrain,  Sex=="male", Age >= 60, Survived==1, Pclass==1)

print(paste("Percentage of male Survivors in Age range GT 60: ", (nrow(maleAgeGT60FclassSurvived)/nrow(maleAgeGT60FclassTot))*100))



#FemaleAgeSclass

maleAge0to15SclassTot <- filter(dfTrain, Sex=="male", Age <= 15, Pclass==2)
maleAge0to15SclassSurvived <- filter(dfTrain,  Sex=="male", Age <= 15, Survived==1, Pclass==2)

print(paste("Percentage of male Survivors in Age range 0 to 15 & Second class: ", (nrow(maleAge0to15SclassTot)/nrow(maleAge0to15SclassTot))*100))




maleAge15to30SclassTot <- filter(dfTrain, Sex=="male", Age > 15 & Age <= 30, Pclass==2)
maleAge15to30SclassSurvived <- filter(dfTrain,  Sex=="male", Age > 15 & Age <= 30, Survived==1, Pclass==2)

print(paste("Percentage of male Survivors in Age range 15 to 30 & second class: ", (nrow(maleAge15to30SclassSurvived)/nrow(maleAge15to30SclassTot))*100))




maleAge30to45SclassTot <- filter(dfTrain, Sex=="male", Age > 30 & Age <= 45, Pclass==2)
maleAge30to45SclassSurvived <- filter(dfTrain,  Sex=="male", Age > 30 & Age <= 45, Survived==1, Pclass==2)

print(paste("Percentage of male Survivors in Age range 30 to 45 & Second class: ", (nrow(maleAge30to45SclassSurvived)/nrow(maleAge30to45SclassTot))*100))



maleAge45to60SclassTot <- filter(dfTrain, Sex=="male", Age > 45 & Age <= 60, Pclass==3)
maleAge45to60SclassSurvived <- filter(dfTrain,  Sex=="male", Age > 45 & Age <= 60, Survived==1, Pclass==3)

print(paste("Percentage of male Survivors in Age range 45 to 60 & Second Class: ", (nrow(maleAge45to60SclassSurvived)/nrow(maleAge45to60SclassTot))*100))




maleAgeGT60SclassTot <- filter(dfTrain, Sex=="male", Age >= 60, Pclass==2)
maleAgeGT60SclassSurvived <- filter(dfTrain,  Sex=="male", Age >= 60, Survived==1, Pclass==2)

print(paste("Percentage of male Survivors in Age range GT 60 & Second class: ", (nrow(maleAgeGT60SclassSurvived)/nrow(maleAgeGT60SclassTot))*100))

#maleAgeTclass

maleAge0to15TclassTot <- filter(dfTrain, Sex=="male", Age <= 15, Pclass==3)
maleAge0to15TclassSurvived <- filter(dfTrain,  Sex=="male", Age <= 15, Survived==1, Pclass==3)

print(paste("Percentage of male Survivors in Age range 0 to 15 & Third class: ", (nrow(maleAge0to15TclassSurvived)/nrow(maleAge0to15TclassTot))*100))




maleAge15to30TclassTot <- filter(dfTrain, Sex=="male", Age > 15 & Age <= 30, Pclass==3)
maleAge15to30TclassSurvived <- filter(dfTrain,  Sex=="male", Age > 15 & Age <= 30, Survived==1, Pclass==3)

print(paste("Percentage of male Survivors in Age range 15 to 30 in Third class: ", (nrow(maleAge15to30TclassSurvived)/nrow(maleAge15to30TclassTot))*100))




maleAge30to45TclassTot <- filter(dfTrain, Sex=="male", Age > 30 & Age <= 45, Pclass==3)
maleAge30to45TclassSurvived <- filter(dfTrain,  Sex=="male", Age > 30 & Age <= 45, Survived==1, Pclass==3)

print(paste("Percentage of male Survivors in Age range 30 to 45 in Third class: ", (nrow(maleAge30to45TclassSurvived)/nrow(maleAge30to45TclassTot))*100))



maleAge45to60TclassTot <- filter(dfTrain, Sex=="male", Age > 45 & Age <= 60, Pclass==3)
maleAge45to60TclassSurvived <- filter(dfTrain,  Sex=="male", Age > 45 & Age <= 60, Survived==1, Pclass==3)

print(paste("Percentage of male Survivors in Age range 45 to 60 in Third class: ", (nrow(maleAge45to60TclassSurvived)/nrow(maleAge45to60TclassTot))*100))




maleAgeGT60TclassTot <- filter(dfTrain, Sex=="male", Age >= 60, Pclass==3)
maleAgeGT60TclassSurvived <- filter(dfTrain,  Sex=="male", Age >= 60, Survived==1, Pclass==3)


print(paste("Percentage of male Survivors in Age range GT 60 in third class: ", (nrow(maleAgeGT60TclassSurvived)/nrow(maleAgeGT60TclassTot))*100))




SurvivalChances <- function(Age, Sex, Pclass)
{
  if(Sex=="female")
  {
    if(Pclass==1)
    {
      print("Chances of survival for a female in First Class: 96%")
      if(Age <= 15)
      {
        print("Chances of survival: 67%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 97%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 100%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 94%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 100%")
      }
    }
    
    else if(Pclass==2)
    {
      print("Chances of survival for a female in Second Class: 92%")
      if(Age <= 15)
      {
        print("Chances of survival: 100%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 91%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 91.3%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 0%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 100%")
      }
    }
    
    else if(Pclass==3)
    {
      print("Chances of survival for a female in Second Class: 50%")
      if(Age <= 15)
      {
        print("Chances of survival: 53%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 49%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 30%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 0%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 100%")
      }
    }
    if(Pclass=="")
    {
      if(Age <= 15)
      {
        print("Chances of survival: 65%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 75%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 78%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 85%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 100%")
      }
    }
  }
  else if(Sex=="male")
  {
    if(Pclass==1)
    {
      print("Chances of survival for a male in First Class: 37%")
      if(Age <= 15)
      {
        print("Chances of survival: 100%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 43%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 48%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 94%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 14%")
      }
    }
    
    else if(Pclass==2)
    {
      print("Chances of survival for a male in Second Class: 16%")
      if(Age <= 15)
      {
        print("Chances of survival: 100%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 2%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 13%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 0%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 25%")
      }
    }
    
    else if(Pclass==3)
    {
      print("Chances of survival for a male in Second Class: 14%")
      if(Age <= 15)
      {
        print("Chances of survival: 32%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 14%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 14%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 0%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 0%")
      }
    }
    if(Pclass=="")
    {
      if(Age <= 15)
      {
        print("Chances of survival: 52.5%")
      }
      else if(Age > 15 & Age <= 30)
      {
        print("Chances of survival: 15%")
      }
      else if(Age>30 & Age <= 45)
      {
        print("Chances of Survival: 22%")
      }
      else if(Age>45 & Age <= 60)
      {
        print("Chances of Survival: 18.5%")
      }
      else if(Age>60)
      {
        print("Chances of Survival: 14%")
      }
    }
  }
}
    

    
SurvivalChances(24, "Male", 1)


```  